\propertitle

### 2019-06-19 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-18T20:14:11+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The context we consider is as follows. We have recorded the spiking
activities of $n$ neurons for a certain amount of time. These neurons are
our \enquote{sample} or \enquote{subnetwork}. Their spikes are binned into
$T$ time bins and binarized to $\set{0,1}$ values in each bin. Call
$\ya_{t}$ the number of neurons that fire during time bin $t$, divided by
the total number of neurons $n$; this is the \emph{normalized total
  activity} of the sample, or just \enquote{activity} for short. Obviously
$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spikes
during bin $t$; if $\ya_{t}=1$, all spike at some point during bin $t$. For
brevity, let's say \enquote{at $t$} for \enquote{during time bin $t$}. From
the activities $\set{\ya_{t}}$ we can count how many times the activity
levels $\ya=0$, $\ya=1/n$, and so on appeared during the recording,
obtaining the measured relative-frequency distribution
$(\yff_{\ya}) \defs \yf$. We can also consider the sample activity at time
bins \emph{outside} of the recorded range. Such activity is unknown to us,
of course.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain regions
\mynote{be more specific}. For many animal species, the neurons that are
recorded within the region are not -- and at present cannot be --
specifically chosen from among the rest, owing to several limiting factors;
for example, limitations in the precision with which electrodes are
inserted, or with which neurons are targeted by viruses. In fact, the set
of recorded neurons may even change during very long recordings or across
experimental sessions. We assume that there's a region, comprising a
network of $N$ neurons, from which other sets of neurons had the same
probability of being recorded as the set of $n$ neurons that was actually
recorded. This is our \enquote{full network}\mynote{how about calling it
  \enquote{the pool}?}. The normalized total activity of these $N$ neurons
at $t$ is $\yA_{t}$. The relative frequencies of the various activity
levels during the recording were $(\yFF_{yA}) \defs \yF$. We don't know the
values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we assume
that we know $N$; in \sect*** we discuss the consequences of our lack of
precise knowledge about this number.

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording. For example: what was its
frequency distribution $\yF$ during the recording? How much total activity
$\yA_{t}$ should we expect at a time bin $t$ during the recording? And at a
new time bin? We cannot answer these questions with certainty; we can only
give distributions of probability over their possible answers. The approach
presented here gives such probability distributions.

@@

@@

We apply the approach just described to two data sets publicly available in
the literature:
\begin{itemize}[wide]
\item The first data set, from \textcite[rat 14147]{stensolaetal2012},
  consists of $n=65$ neurons (27 of which classified as grid cells) from
  rat Medial Entorhinal Cortex, recorded for about 20 minutes. Their spikes
  are binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
\item The second data set, from
  \textcite{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
  consists of $n=159$ neurons from macaque Motor Cortex, recorded for about
  15 minutes. Their spikes are binned into $T=300\,394$ bins of
  $3\;\textrm{ms}$ width.
\end{itemize}

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-19T11:23:16+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The context we consider is as follows. During an experimental session we
have recorded the spiking activities of $n$ neurons for a certain amount of
time. These neurons are our \enquote{sample} or \enquote{subnetwork}. Their
spikes are binned into $T$ time bins and binarized to $\set{0,1}$ values in
each bin. Call $\ya_{t}$ the number of neurons that fire during time bin
$t$, divided by the total number of neurons $n$; this is the
\emph{normalized total activity} of the sample, or just \enquote{activity}
for short. Obviously $\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if
$\ya_{t}=0$, no neuron spikes during bin $t$; if $\ya_{t}=1$, all spike at
some point during bin $t$. For brevity, let's say \enquote{at $t$} for
\enquote{during time bin $t$}. From the activities $\set{\ya_{t}}$ we can
count how many times the activity levels $\ya=0$, $\ya=1/n$, and so on
appeared during the recording, obtaining the measured relative-frequency
distribution $(\yff_{\ya}) \defs \yf$. We can also consider the sample
activity at time bins \emph{outside} of the recorded range. Such activity
is unknown to us, of course.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain regions
\mynote{be more specific}. For many animal species, the neurons that are
recorded within the region are not -- and at present cannot be --
specifically chosen from among the rest, owing to several limiting factors;
for example, limitations in the precision with which electrodes are
inserted, or with which neurons are targeted by viruses. In fact, the set
of recorded neurons may even change during very long recordings or across
experimental sessions. We assume that there's a region, comprising a
network of $N$ neurons, from which other sets of neurons had or will have
the same probability of being recorded, even in other experimental
sessions, as the set of $n$ neurons that was actually recorded in this
session. This is our \enquote{full network}\mynote{how about calling it
  \enquote{the pool}?}. The normalized total activity of these $N$ neurons
at $t$ is $\yA_{t}$. The relative frequencies of the various activity
levels during the recording were $(\yFF_{\yA}) \defs \yF$. We don't know the
values $\yA_{t}$ at each $t$, or the frequency distribution $\yF$. We only
know for certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the time being we assume
that we know $N$; in \sect*** we discuss the consequences of our lack of
precise knowledge about this number.

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording, and across sessions under the
same study conditions. For example: what was its frequency distribution
during the recording? How much does this frequency distribution change
across sessions? How much total activity should we expect at any time bin
during a recording? We cannot answer these questions with certainty; we can
only give distributions of probability over their possible answers. The
approach presented here gives such probability distributions.

\mynote{Move this to intro?}We want to stress the usefulness of making
inferences about the full-network activity. First of all, this seems to be
the primary idea behind recording a sample. Second, it allows us to make
comparisons across experimental sessions; such comparisons would me
difficult or meaningless made at the level of recorded samples, which
generally comprise non-overlapping sets of neurons differing also in size.

@@

The calculation amounts to a convex optimization
\citep{meadetal1984}[\chap~10]{pressetal1988_r2007}{fangetal1997,boydetal2004_r2009,portamana2017b}
and for the numbers $N,n,m$ considered in the next section it can be done
on a modern computer without approximations of normalization constants or
potential functions.

@@

We apply the approach just described to two data sets publicly available in
the literature:
\begin{itemize}[wide]
\item The first, from \textcite[rat 14147]{stensolaetal2012},
  consists of $n=65$ neurons (27 of which classified as grid cells) from
  rat Medial Entorhinal Cortex, recorded for about 20 minutes. Their spikes
  are binned into $T=417\,641$ bins of $3\;\textrm{ms}$ width.
\item The second, from
  \textcite{rostamietal2016_r2017}, %\texttt{data\_RS.npy}
  consists of $n=159$ neurons from macaque Motor Cortex, recorded for about
  15 minutes. Their spikes are binned into $T=300\,394$ bins of
  $3\;\textrm{ms}$ width.
\end{itemize}

We first calculate the maximum-entropy distribution for the full network by
using six moments. This number already provides as much information as the
full frequency distribution of the sample, and at the same time illustrates
the use of the approach in questions of statistic sufficiency (typically
limited to two or three moments). We show the results for full-network
sizes $N=1\,000$, $N=5\,000$, $N=10\,000$\mynote{motivate?}. Thereafter we
show the results for two moments and six moments using $N=10\,000$.

\begin{figure}[!b]
\centering
\includegraphics[width=\linewidth]{scripts/paper_stensola_6mom_distributionsN_prior2.pdf}%
\\%
\includegraphics[width=\linewidth]{scripts/paper_riehle_6mom_distributionsN_prior2.pdf}%
\caption{***}
\label{fig:plots_6mom}
\end{figure}
%maxent_solutions_different_moments_prior2_6moments.nb
